Metropolis-Hastings Implementations of the Forward-Filtering-Backward-Sampling \\ Algorithm for Particle Smoothing

### 2013-11-28 Paper/complete/preprint_MCMCsmoother.tex
Added clean version of code.

\documentclass[peerreview,11pt,draftcls,onecolumn]{IEEEtran}

%% Packages
\usepackage{cite}
\usepackage[pdftex]{graphicx} \graphicspath{{images/}}
\usepackage[cmex10]{amsmath}
%\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{array}
\usepackage{mdwmath}
\usepackage{mdwtab}
\usepackage{eqparbox}
\usepackage[caption=false,font=footnotesize]{subfig}
\usepackage{fixltx2e}
%\usepackage{stfloats}
\usepackage{url}
\usepackage{color}

@@

\author{Pete~Bunch*,~\IEEEmembership{Member,~IEEE,} and
        Simon~Godsill,~\IEEEmembership{Member,~IEEE}% <-this % stops a space
\thanks{P. Bunch and S. Godsill are with the Department
of Engineering, Cambridge University, UK. email: \{pb404,sjg30\}@cam.ac.uk}% <-this % stops a space
\thanks{Manuscript received April 03, 2012.}}

@@

%\begin{IEEEkeywords}
%state space model, particle filter, smoothing, MCMC, Bayesian inference
%\end{IEEEkeywords}

\ifCLASSOPTIONpeerreview
\begin{center} \bfseries EDICS Categories: MLR-BAYL, SSP-NGAU, SSP-TRAC, SSP-FILT \end{center}
\fi

\newpage
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
%\IEEEpeerreviewmaketitle

@@

\IEEEPARstart{T}{he} objective of sequential Bayesian inference is the estimation of an unknown, time-varying quantity from incomplete or inaccurate observations. This is commonly achieved through the use of discrete-time probabilistic models for the evolution of the latent state and the measurement process, defined by transition and observation densities,
%
\begin{IEEEeqnarray}{rCl}
x_{k} & \sim & p(x_{k}|x_{k-1}) \label{eq:state_proc}\\
y_{k} & \sim & p(y_{k}|x_{k})   \label{eq:observ_proc}   .
\end{IEEEeqnarray}

@@

\begin{figure}[!t]
\centering
\subfloat[Case 1]{\includegraphics[width=0.45\columnwidth]{case1_trajectory.pdf}}
\\
\subfloat[Case 2]{\includegraphics[width=0.45\columnwidth]{case2_trajectory.pdf}}
\\
\subfloat[Case 3]{\includegraphics[width=0.45\columnwidth]{case3_trajectory.pdf}}
\caption{Example trajectory with observations from the two noise covariance cases. Constant likelihood contours are shown for the first and last position (dashed)}
\label{fig:example_trajectories}
\end{figure}

@@

\begin{table}[!t]%
\renewcommand{\arraystretch}{1.3}
\caption{Smoother performance measures for noise covariance case 1. All results are averaged over 100 runs, with 500 time steps in each run.}
\label{tab:case1_performance}
{\scriptsize
\begin{center}
\begin{tabular}{|c||c|c|c|c|c|}
\hline
Algorithm & \begin{minipage}[c]{0.9cm} Position RMSE \end{minipage} & \begin{minipage}[c]{0.9cm}  Velocity RMSE \end{minipage} & \begin{minipage}[c]{0.9cm}  ENEES \end{minipage} & \begin{minipage}[c]{0.9cm}  Running Time (s) \end{minipage} & \begin{minipage}[c]{0.9cm}  Unique Particles \end{minipage} \\
\hline
FS             & 0.56 & 0.96 & 0.99 & 0.00 & 2.13 \\
D-FFBS         & 0.45 & 0.76 & 0.84 & 65.88 & 20.54 \\
\hline
MH-FFBS (1)    & 0.48 & 0.81 & 0.90 & 1.80 & 13.95 \\
MH-FFBS (3)    & 0.46 & 0.78 & 0.87 & 3.89 & 17.63 \\
MH-FFBS (10)   & 0.46 & 0.77 & 0.85 & 11.20 & 19.83 \\
MH-FFBS (30)   & 0.45 & 0.76 & 0.84 & 32.05 & 20.40 \\
MH-FFBS (100)  & 0.45 & 0.76 & 0.84 & 105.12 & 20.55 \\
\hline
MH-FFBP (1)    & 0.45 & 0.75 & 0.82 & 9.20 & 44.32 \\
MH-FFBP (3)    & 0.43 & 0.73 & 0.79 & 19.23 & 70.12 \\
MH-FFBP (10)   & 0.43 & 0.72 & 0.78 & 54.33 & 90.10 \\
MH-FFBP (30)   & 0.42 & 0.71 & 0.79 & 154.47 & 96.89 \\
MH-FFBP (100)  & 0.42 & 0.70 & 0.79 & 504.58 & 98.53 \\
\hline
\end{tabular}
\end{center}
}
\end{table}

\begin{table}[!t]%
\renewcommand{\arraystretch}{1.3}
\caption{Smoother performance measures for noise covariance case 2. All results are averaged over 100 runs, with 500 time steps in each run.}
\label{tab:case2_performance}
{\scriptsize
\begin{center}
\begin{tabular}{|c||c|c|c|c|c|}
\hline
Algorithm & \begin{minipage}[c]{0.9cm} Position RMSE \end{minipage} & \begin{minipage}[c]{0.9cm}  Velocity RMSE \end{minipage} & \begin{minipage}[c]{0.9cm}  ENEES \end{minipage} & \begin{minipage}[c]{0.9cm}  Running Time (s) \end{minipage} & \begin{minipage}[c]{0.9cm}  Unique Particles \end{minipage} \\
\hline
FS             & 8.01 & 1.90 & 0.99 & 0.00 & 2.75 \\
D-FFBS         & 7.71 & 1.70 & 0.92 & 66.40 & 14.06 \\
\hline
MH-FFBS (1)    & 7.91 & 1.83 & 0.98 & 1.79 & 6.51 \\
MH-FFBS (3)    & 7.83 & 1.78 & 0.96 & 3.86 & 8.63 \\
MH-FFBS (10)   & 7.77 & 1.73 & 0.93 & 11.09 & 11.39 \\
MH-FFBS (30)   & 7.75 & 1.71 & 0.92 & 31.77 & 13.14 \\
MH-FFBS (100)  & 7.70 & 1.70 & 0.92 & 104.14 & 13.88 \\
\hline
MH-FFBP (1)    & 7.75 & 1.72 & 0.92 & 9.25 & 22.97 \\
MH-FFBP (3)    & 7.67 & 1.67 & 0.88 & 19.34 & 42.22 \\
MH-FFBP (10)   & 7.62 & 1.63 & 0.85 & 54.61 & 70.06 \\
MH-FFBP (30)   & 7.60 & 1.61 & 0.84 & 155.38 & 88.81 \\
MH-FFBP (100)  & 7.59 & 1.60 & 0.84 & 507.58 & 96.82 \\
\hline
\end{tabular}
\end{center}
}
\end{table}

\begin{table}[!t]%
\renewcommand{\arraystretch}{1.3}
\caption{Smoother performance measures for noise covariance case 3. All results are averaged over 100 runs, with 500 time steps in each run.}
\label{tab:case3_performance}
{\scriptsize
\begin{center}
\begin{tabular}{|c||c|c|c|c|c|}
\hline
Algorithm & \begin{minipage}[c]{0.9cm} Position RMSE \end{minipage} & \begin{minipage}[c]{0.9cm}  Velocity RMSE \end{minipage} & \begin{minipage}[c]{0.9cm}  ENEES \end{minipage} & \begin{minipage}[c]{0.9cm}  Running Time (s) \end{minipage} & \begin{minipage}[c]{0.9cm}  Unique Particles \end{minipage} \\
\hline
FS            & 7.41 & 2.15 & 0.98 & 0.00 & 3.97 \\
D-FFBS        & 7.19 & 2.03 & 0.97 & 66.27 & 7.48 \\
\hline
MH-FFBS (1)   & 7.39 & 2.14 & 0.98 & 1.75 & 4.79 \\
MH-FFBS (3)   & 7.36 & 2.12 & 0.98 & 3.75 & 5.33 \\
MH-FFBS (10)  & 7.30 & 2.08 & 0.98 & 10.78 & 6.16 \\
MH-FFBS (30)  & 7.24 & 2.05 & 0.97 & 30.83 & 6.88 \\
MH-FFBS (100) & 7.20 & 2.04 & 0.97 & 101.02 & 7.38 \\
\hline
MH-FFBP (1)   & 7.22 & 2.05 & 0.97 & 9.12 & 12.57 \\
MH-FFBP (3)   & 7.08 & 1.98 & 0.94 & 19.01 & 22.78 \\
MH-FFBP (10)  & 6.95 & 1.92 & 0.92 & 53.64 & 44.15 \\
MH-FFBP (30)  & 6.87 & 1.88 & 0.90 & 152.42 & 70.71 \\
MH-FFBP (100) & 6.84 & 1.86 & 0.88 & 498.38 & 91.35 \\
\hline
\end{tabular}
\end{center}
}
\end{table}

\begin{figure}[!t]
\centering
\subfloat[Case 1]{\includegraphics[width=0.45\columnwidth]{case1_smoother_comparison_posRMSE_time.pdf}}
\\
\subfloat[Case 2]{\includegraphics[width=0.45\columnwidth]{case2_smoother_comparison_posRMSE_time.pdf}}
\\
\subfloat[Case 3]{\includegraphics[width=0.45\columnwidth]{case3_smoother_comparison_posRMSE_time.pdf}}
\caption{Position RMSEs and running times of various smoothing algorithms with different noise covariance cases.}%
\label{fig:rmse_vs_time}%
\end{figure}

>>>>

\documentclass{article}

%% Packages
\usepackage{IEEEtrantools}
\usepackage{cite}
\usepackage[pdftex]{graphicx} \graphicspath{{images/}}
\usepackage[cmex10]{amsmath}
%\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{array}
\usepackage{mdwmath}
\usepackage{mdwtab}
\usepackage{eqparbox}
\usepackage[caption=false,font=footnotesize]{subfig}
\usepackage{fixltx2e}
%\usepackage{stfloats}
\usepackage{url}
\usepackage{color}

@@

\author{Pete~Bunch and Simon~Godsill% <-this % stops a space
\thanks{Department of Engineering, Cambridge University, UK. email: \{pb404,sjg30\}@cam.ac.uk}}

@@

@@

The objective of sequential Bayesian inference is the estimation of an unknown, time-varying quantity from incomplete or inaccurate observations. This is commonly achieved through the use of discrete-time probabilistic models for the evolution of the latent state and the measurement process, defined by transition and observation densities,
%
\begin{IEEEeqnarray}{rCl}
x_{k} & \sim & p(x_{k}|x_{k-1}) \label{eq:state_proc}\\
y_{k} & \sim & p(y_{k}|x_{k})   \label{eq:observ_proc}   .
\end{IEEEeqnarray}

@@

\begin{figure}[!t]
\centering
\subfloat[Case 1]{\includegraphics[width=0.4\columnwidth]{case1_trajectory.pdf}}
\\
\subfloat[Case 2]{\includegraphics[width=0.4\columnwidth]{case2_trajectory.pdf}}
\\
\subfloat[Case 3]{\includegraphics[width=0.4\columnwidth]{case3_trajectory.pdf}}
\caption{Example trajectory with observations from the two noise covariance cases. Constant likelihood contours are shown for the first and last position (dashed)}
\label{fig:example_trajectories}
\end{figure}

@@

\begin{table}[!t]%
\renewcommand{\arraystretch}{1.3}
\caption{Smoother performance measures for noise covariance case 1. All results are averaged over 100 runs, with 500 time steps in each run.}
\label{tab:case1_performance}
{\scriptsize
\centering
\begin{tabular}{|c||c|c|c|c|c|}
\hline
Algorithm & \begin{minipage}[c]{0.9cm} Position RMSE \end{minipage} & \begin{minipage}[c]{0.9cm}  Velocity RMSE \end{minipage} & \begin{minipage}[c]{0.9cm}  ENEES \end{minipage} & \begin{minipage}[c]{0.9cm}  Running Time (s) \end{minipage} & \begin{minipage}[c]{0.9cm}  Unique Particles \end{minipage} \\
\hline
FS             & 0.56 & 0.96 & 0.99 & 0.00 & 2.13 \\
D-FFBS         & 0.45 & 0.76 & 0.84 & 65.88 & 20.54 \\
\hline
MH-FFBS (1)    & 0.48 & 0.81 & 0.90 & 1.80 & 13.95 \\
MH-FFBS (3)    & 0.46 & 0.78 & 0.87 & 3.89 & 17.63 \\
MH-FFBS (10)   & 0.46 & 0.77 & 0.85 & 11.20 & 19.83 \\
MH-FFBS (30)   & 0.45 & 0.76 & 0.84 & 32.05 & 20.40 \\
MH-FFBS (100)  & 0.45 & 0.76 & 0.84 & 105.12 & 20.55 \\
\hline
MH-FFBP (1)    & 0.45 & 0.75 & 0.82 & 9.20 & 44.32 \\
MH-FFBP (3)    & 0.43 & 0.73 & 0.79 & 19.23 & 70.12 \\
MH-FFBP (10)   & 0.43 & 0.72 & 0.78 & 54.33 & 90.10 \\
MH-FFBP (30)   & 0.42 & 0.71 & 0.79 & 154.47 & 96.89 \\
MH-FFBP (100)  & 0.42 & 0.70 & 0.79 & 504.58 & 98.53 \\
\hline

\end{tabular}
}
\end{table}

\begin{table}[!t]%
\renewcommand{\arraystretch}{1.3}
\caption{Smoother performance measures for noise covariance case 2. All results are averaged over 100 runs, with 500 time steps in each run.}
\label{tab:case2_performance}
{\scriptsize
\centering
\begin{tabular}{|c||c|c|c|c|c|}
\hline
Algorithm & \begin{minipage}[c]{0.9cm} Position RMSE \end{minipage} & \begin{minipage}[c]{0.9cm}  Velocity RMSE \end{minipage} & \begin{minipage}[c]{0.9cm}  ENEES \end{minipage} & \begin{minipage}[c]{0.9cm}  Running Time (s) \end{minipage} & \begin{minipage}[c]{0.9cm}  Unique Particles \end{minipage} \\
\hline
FS             & 8.01 & 1.90 & 0.99 & 0.00 & 2.75 \\
D-FFBS         & 7.71 & 1.70 & 0.92 & 66.40 & 14.06 \\
\hline
MH-FFBS (1)    & 7.91 & 1.83 & 0.98 & 1.79 & 6.51 \\
MH-FFBS (3)    & 7.83 & 1.78 & 0.96 & 3.86 & 8.63 \\
MH-FFBS (10)   & 7.77 & 1.73 & 0.93 & 11.09 & 11.39 \\
MH-FFBS (30)   & 7.75 & 1.71 & 0.92 & 31.77 & 13.14 \\
MH-FFBS (100)  & 7.70 & 1.70 & 0.92 & 104.14 & 13.88 \\
\hline
MH-FFBP (1)    & 7.75 & 1.72 & 0.92 & 9.25 & 22.97 \\
MH-FFBP (3)    & 7.67 & 1.67 & 0.88 & 19.34 & 42.22 \\
MH-FFBP (10)   & 7.62 & 1.63 & 0.85 & 54.61 & 70.06 \\
MH-FFBP (30)   & 7.60 & 1.61 & 0.84 & 155.38 & 88.81 \\
MH-FFBP (100)  & 7.59 & 1.60 & 0.84 & 507.58 & 96.82 \\
\hline
\end{tabular}
}
\end{table}

\begin{table}[!t]%
\renewcommand{\arraystretch}{1.3}
\caption{Smoother performance measures for noise covariance case 3. All results are averaged over 100 runs, with 500 time steps in each run.}
\label{tab:case3_performance}
{\scriptsize
\centering
\begin{tabular}{|c||c|c|c|c|c|}
\hline
Algorithm & \begin{minipage}[c]{0.9cm} Position RMSE \end{minipage} & \begin{minipage}[c]{0.9cm}  Velocity RMSE \end{minipage} & \begin{minipage}[c]{0.9cm}  ENEES \end{minipage} & \begin{minipage}[c]{0.9cm}  Running Time (s) \end{minipage} & \begin{minipage}[c]{0.9cm}  Unique Particles \end{minipage} \\
\hline
FS            & 7.41 & 2.15 & 0.98 & 0.00 & 3.97 \\
D-FFBS        & 7.19 & 2.03 & 0.97 & 66.27 & 7.48 \\
\hline
MH-FFBS (1)   & 7.39 & 2.14 & 0.98 & 1.75 & 4.79 \\
MH-FFBS (3)   & 7.36 & 2.12 & 0.98 & 3.75 & 5.33 \\
MH-FFBS (10)  & 7.30 & 2.08 & 0.98 & 10.78 & 6.16 \\
MH-FFBS (30)  & 7.24 & 2.05 & 0.97 & 30.83 & 6.88 \\
MH-FFBS (100) & 7.20 & 2.04 & 0.97 & 101.02 & 7.38 \\
\hline
MH-FFBP (1)   & 7.22 & 2.05 & 0.97 & 9.12 & 12.57 \\
MH-FFBP (3)   & 7.08 & 1.98 & 0.94 & 19.01 & 22.78 \\
MH-FFBP (10)  & 6.95 & 1.92 & 0.92 & 53.64 & 44.15 \\
MH-FFBP (30)  & 6.87 & 1.88 & 0.90 & 152.42 & 70.71 \\
MH-FFBP (100) & 6.84 & 1.86 & 0.88 & 498.38 & 91.35 \\
\hline
\end{tabular}
}
\end{table}

\begin{figure}[!t]
\centering
\subfloat[Case 1]{\includegraphics[width=0.4\columnwidth]{case1_smoother_comparison_posRMSE_time.pdf}}
\\
\subfloat[Case 2]{\includegraphics[width=0.4\columnwidth]{case2_smoother_comparison_posRMSE_time.pdf}}
\\
\subfloat[Case 3]{\includegraphics[width=0.4\columnwidth]{case3_smoother_comparison_posRMSE_time.pdf}}
\caption{Position RMSEs and running times of various smoothing algorithms with different noise covariance cases.}%
\label{fig:rmse_vs_time}%
\end{figure}

### 2013-11-28 Poster/MCMC_smoothing_poster.tex
Added clean version of code.

\begin{abstract}
The conventional forward-filtering-backward-sampling algorithm for particle smoothing generates samples from the joint smoothing distribution by sequentially sampling backwards from the forward filter approximations. This algorithm has a computational complexity quadratic in the number of particles, which is often restrictive. In addition, the support of the smoothing distribution is restricted to the values which appear in the filtering approximation. Here a Metropolis-Hastings sampling procedure is used to improve the efficiency of the forward-filtering-backward-sampling algorithm, achieving comparable error performance but with a lower execution time. In addition, an algorithm for approximating the joint smoothing distribution without limited support is presented, which achieves simultaneous improvements in both execution time and error. These algorithms also provide a greater degree of flexibility over existing methods, allowing a trade-off between execution time and estimation error, controlled by the length of the Markov chains.
\end{abstract}

@@

@@

Consider a standard hidden Markov model for Bayesian inference,
%
\begin{IEEEeqnarray*}{rCl}
x_{k} & \sim & p(x_{k}|x_{k-1}) \\
y_{k} & \sim & p(y_{k}|x_{k})      .
\end{IEEEeqnarray*}

@@

For non-linear, non-Gaussian models, numerical approximations are needed. Particle methods represent posterior distributions with a set of samples \cite{Doucet2009}.
\begin{itemize}
  \item Filter: $\hat{p}(x_{k}|y_{1:k}) = \sum_i w_k^{(i)} \delta_{x_k^{(i)}}(x_k)$.
  \item Smoother: $\hat{p}(x_{1:K}|y_{1:K}) = \sum_i w_K^{(i)} \delta_{x_{1:K}^{(i)}}(x_{1:K})$.
\end{itemize}

@@

This may be sampled easily by calculating $\tilde{w}_k^{(i)}$ for each of the $N_F$ filter particles and sampling the resulting multinomial distribution. The whole procedure is repeated $N_S$ times, once for each smoothing particle.

\begin{algorithm}
\begin{algorithmic}[1]
\linespread{1.7} \selectfont
 	\STATE Run a particle filter to approximate $p(x_k|y_{1:k})$ for each $k$ and store the resulting particles $\{x_k^{(i)}, w_k^{(i)}\}$.
	\FOR{$i = 1$ \TO $N_S$}
		\STATE Sample $\tilde{x}_{K}^{(i)} \sim \hat{p}(x_K|y_{1:K})$.
		\FOR{$k = K-1$ \TO $1$}
            \STATE Sample $\tilde{x}_{k}^{(i)} \sim p(x_k|\tilde{x}_{k+1}, y_{1:K})$. This is the critical step. See text for various methods.
%			\FOR{$j = 1$ \TO $N_F$}
%				\STATE Calculate weight $\tilde{w}_k^{(j)} = w_k^{(j)} p(\tilde{x}_{k+1}|x_k^{(j)})$.
%			\ENDFOR
%			\STATE Sample $\tilde{x}_{k}^{(i)} \sim \sum_j \tilde{w}_k^{(j)} \delta_{x_{k}^{(j)}}(x_{k})$.
		\ENDFOR
	\ENDFOR
\end{algorithmic}
\caption{Forward-filtering-backward-sampling algorithm}
\end{algorithm}

@@

Modify the MH proposal to use,
%
\begin{IEEEeqnarray*}{c}
\hat{p}(x_{1:k-1}|y_{1:k-1}) q(x_k|x_{k-1}, \tilde{x}_{k+1}, y_k)     .
\end{IEEEeqnarray*}

Support of $x_k$ is no longer restricted to those values appearing in the filter approximation. This is inspired by the method used in \cite{Fearnhead2010} for two-filter marginal smoothing.

Acceptance probability is now,
%
\begin{IEEEeqnarray*}{rCl}
\alpha &=& \min \Bigg[ 1, \frac{  p(\tilde{x}_{k+1}|x_k^{*}) p(x_k^{*}|x_{k-1}^{*}) p(y_k|x_k^{*}) q(x_k|x_{k-1},\tilde{x}_{k+1},y_k) }{ p(\tilde{x}_{k+1}|x_k) p(x_k|x_{k-1}) p(y_k|x_k) q(x_k^{*}|x_{k-1}^{*},\tilde{x}_{k+1},y_k) }  \Bigg]      .
\end{IEEEeqnarray*}

>>>>

\begin{abstract}
The conventional forward-filtering-backward-sampling algorithm for particle smoothing generates samples from the joint smoothing distribution by sequentially sampling backwards from the forward filter approximations. This algorithm has a computational complexity quadratic in the number of particles, which is often restrictive. In addition, the support of the smoothing distribution is restricted to the values which appear in the filtering approximation. Here a Metropolis-Hastings sampling procedure is used to improve the efficiency of the forward-filtering-backward-sampling algorithm. In addition, an algorithm for approximating the joint smoothing distribution without limited support is presented, which achieves simultaneous improvements in both execution time and error. These algorithms also provide a greater degree of flexibility over existing methods, allowing a trade-off between execution time and estimation error, controlled by the length of the Markov chains.
\end{abstract}

@@

\begingroup
{\small
\renewcommand{\section}[2]{}%
To be published as:
\begin{thebibliography}
\newblock P.~Bunch and S.~Godsill.
\newblock ``Improved particle approximations to the joint smoothing distribution using Markov chain Monte Carlo.''
\newblock {\em IEEE Transactions on Signal Processing}.
\end{thebibliography}
}
\endgroup

@@

Consider a hidden Markov model for Bayesian inference,
%
\begin{IEEEeqnarray*}{rCl}
x_{k} & \sim & p(x_{k}|x_{k-1}) \\
y_{k} & \sim & p(y_{k}|x_{k})      .
\end{IEEEeqnarray*}

@@

%For non-linear, non-Gaussian models, numerical approximations are needed. Particle methods represent posterior distributions with a set of samples \cite{Doucet2009}.
For non-linear, non-Gaussian models, particle methods may be used to represent posterior distributions with a set of samples \cite{Doucet2009}.
\begin{itemize}
  \item Filter: $\hat{p}(x_{k}|y_{1:k}) = \sum_i w_k^{(i)} \delta_{x_k^{(i)}}(x_k)$.
  \item Smoother: $\hat{p}(x_{1:K}|y_{1:K}) = \sum_i w_K^{(i)} \delta_{x_{1:K}^{(i)}}(x_{1:K})$.
\end{itemize}

@@

This may be sampled easily by calculating $\tilde{w}_k^{(i)}$ for each of the $N_F$ filter particles and sampling the resulting categorical distribution. The whole procedure is repeated $N_S$ times, once for each smoothing particle.

\begin{algorithm}
\begin{algorithmic}[1]
\linespread{1.7} \selectfont
 	\STATE Run a particle filter to approximate $p(x_k|y_{1:k})$ for each $k$ and store the resulting particles $\{x_k^{(i)}, w_k^{(i)}\}$.
	\FOR{$i = 1$ \TO $N_S$}
		\STATE Sample $\tilde{x}_{K}^{(i)} \sim \hat{p}(x_K|y_{1:K})$.
		\FOR{$k = K-1$ \TO $1$}
            \STATE Sample $\tilde{x}_{k}^{(i)} \sim p(x_k|\tilde{x}_{k+1}, y_{1:K})$. \\
            \emph{This is the critical step. See text for various methods.}
%			\FOR{$j = 1$ \TO $N_F$}
%				\STATE Calculate weight $\tilde{w}_k^{(j)} = w_k^{(j)} p(\tilde{x}_{k+1}|x_k^{(j)})$.
%			\ENDFOR
%			\STATE Sample $\tilde{x}_{k}^{(i)} \sim \sum_j \tilde{w}_k^{(j)} \delta_{x_{k}^{(j)}}(x_{k})$.
		\ENDFOR
	\ENDFOR
\end{algorithmic}
\caption{Forward-filtering-backward-sampling algorithm}
\end{algorithm}

@@

Modify the MH proposal to use,
%
\begin{IEEEeqnarray*}{rCl}
\{ x_{1:k-1}^{(i^*)}, x_k^*  \} &\sim& \hat{p}(x_{1:k-1}|y_{1:k-1}) q(x_k|x_{k-1}, \tilde{x}_{k+1}, y_k)     .
\end{IEEEeqnarray*}

Support of $x_k$ is no longer restricted to those values appearing in the filter approximation (although support of $x_{1:k-1}$ is). This is inspired by the method used in \cite{Fearnhead2010} for two-filter marginal smoothing.

Acceptance probability is now,
%
\begin{IEEEeqnarray*}{rCl}
\alpha &=& \min \Bigg[ 1, \frac{  p(\tilde{x}_{k+1}|x_k^{*}) p(x_k^{*}|x_{k-1}^{(i^*)}) p(y_k|x_k^{*}) q(x_k|x_{k-1},\tilde{x}_{k+1},y_k) }{ p(\tilde{x}_{k+1}|x_k) p(x_k|x_{k-1}) p(y_k|x_k) q(x_k^{*}|x_{k-1}^{(i^*)},\tilde{x}_{k+1},y_k) }  \Bigg]      .
\end{IEEEeqnarray*}
